Music Perception: Pilot Accuracy Analysis

### 2021-01-25 analysis/analysis.Rmd
update analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Perception: Pilot Analysis"
author: "Atsuko Tominaga"
output: html_notebook
---

@@

I hope these make sense...

@@

```{r read, include = FALSE}
# read files and combine them
data_ls <- list.files("data")
combined <- lapply(data_ls, function(f){
  fread(paste("data/", f, sep = ""), sep = ",")
})
data <- do.call(rbind.data.frame, combined)
```

# Results
## Individual data
```{r summary, echo = FALSE}
# remove irrelevant strings
data$midFile <- gsub("./practice/", "", data$midFile)
data$midFile <- gsub("./mid/", "", data$midFile)
data$midFile <- gsub(".mid", "", data$midFile)

# name conditions
data$condition <- "NA"
data$technique <- "NA"
data[grepl("_t_", midFile)]$condition <- "teaching"
data[grepl("_p_", midFile)]$condition <- "performing"
data[grepl("a_", midFile)]$technique <- "articulation"
data[grepl("d_", midFile)]$technique <- "dynamics"
data$item <- gsub("a_t_", "", data$midFile)
data$item <- gsub("d_t_", "", data$item)
data$item <- gsub("a_p_", "", data$item)
data$item <- gsub("d_p_", "", data$item)

# check correct or incorrect
data$correct <- 0
data[condition == "teaching" & rating == "Yes"]$correct <- 1
data[condition == "performing" & rating == "No"]$correct <- 1

all <- data[expMode == "experiment", .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique)]
all[order(subjectNumber, condition, technique)]
```

## 1. Summary
### all
```{r acc, echo = FALSE}
# individual
ind_data <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

@@

- Surprisingly, it showed a trend ( p = .06) but the 95% confidence interval overlaps with 50 %.

### articulation vs. dynamics

```{r plot2, echo = FALSE}
ind_data_ad <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, technique)]
ind_data_ad <- ind_data_ad[order(subjectNumber, technique)]

subset_ad <- ind_data_ad[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/48)*100, SD_Percent = (sd(Sum)/48)*100), by = .(technique)]
subset_ad

data_box <- data.table(subjectNumber = ind_data_ad$subjectNumber, articulation = ind_data_ad[technique == "articulation"]$Percent, dynamics = ind_data_ad[technique == "dynamics"]$Percent)
box <- ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box
```

## One-sample *t*-test
### articulation
```{r t-test-a, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[technique == "articulation"]$Percent)

t.test(ind_data_ad[technique == "articulation"]$Percent, mu = 50, alternative = "two.sided")
```
- Not significantly different from chance (p > .2).

### dynamics
```{r t-test-d, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[technique == "dynamics"]$Percent)

t.test(ind_data_ad[technique == "dynamics"]$Percent, mu = 50, alternative = "two.sided")
```

- The performance was significantly better than chance (p < .01).

## Paired *t*-test between articulation and dynamics
```{r p-t-test, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[technique == "dynamics"]$Percent-ind_data_ad[technique == "articulation"]$Percent)

t.test(ind_data_ad[technique == "articulation"]$Percent, ind_data_ad[technique == "dynamics"]$Percent, paired = TRUE)
```

- There was a difference between articulation and dynamics in terms of accuracy.

## 2. teaching (each item)
### all
```{r item1, echo = FALSE}
# individual
ind_data3 <- data[expMode == "experiment" & subjectNumber != 9 & condition == "teaching", .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique, item)]
ind_data3 <- ind_data3[order(subjectNumber, condition, technique, item)]

# all
subset3 <- ind_data3[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(condition, technique, item)]
subset3[order(-Mean_Percent)]
```

### articulation
```{r item1-a, echo = FALSE}
subset3_a <- subset3[technique == "articulation"]
subset3_a[order(-Mean_Percent)]

bar_a <- ggplot(subset3[technique == "articulation"], aes(x = reorder(item, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar_a
```

### dynamics
```{r item1-d, echo = FALSE}
subset3_d <- subset3[order(-Mean_Percent) & technique == "dynamics"]
subset3_d[order(-Mean_Percent)]

bar_d <- ggplot(subset3[technique == "dynamics"], aes(x = reorder(item, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar_d
```

## 3. performing (each item)
### all
```{r item2, echo = FALSE}
# individual
ind_data4 <- data[expMode == "experiment" & subjectNumber != 9 & condition == "performing", .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique, item)]
ind_data4 <- ind_data4[order(subjectNumber, condition, technique, item)]

# all
subset4 <- ind_data4[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(condition, technique, item)]
subset4[order(-Mean_Percent)]
```

### articulation
```{r item2-a, echo = FALSE}
subset4_a <- subset4[technique == "articulation"]
subset4_a[order(-Mean_Percent)]

bar2_a <- ggplot(subset4[technique == "articulation"], aes(x = reorder(item, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar2_a
```

### dynamics
```{r item2-d, echo = FALSE}
subset4_d <- subset4[technique == "dynamics"]
subset4_d[order(-Mean_Percent)]

bar2_d <- ggplot(subset4[technique == "dynamics"], aes(x = reorder(item, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar2_d
```

```{r export, include = FALSE}
order_data <- rbind(
  data.table(subset3_a$condition, subset3_a$technique, subset3_a$item, subset3_a$Mean_Percent),
  data.table(subset3_d$condition, subset3_d$technique, subset3_d$item, subset3_d$Mean_Percent),
  data.table(subset4_a$condition, subset4_a$technique, subset4_a$item, subset4_a$Mean_Percent),
  data.table(subset4_d$condition, subset4_d$technique, subset4_d$item, subset4_d$Mean_Percent))

colnames(order_data) <- c("condition", "technique", "item", "mean_percent")

>>>>

---
title: "Music Perception: Pilot Accuracy Analysis"
author: "Atsuko Tominaga"
output: html_notebook
---

@@

@@

```{r read, include = FALSE}
# read files and combine them
data_ls <- list.files("data")
combined <- lapply(data_ls, function(f){
  fread(paste("data/", f, sep = ""), sep = ",")
})
all_data <- do.call(rbind.data.frame, combined)
```

# Results
## Individual data
```{r summary, echo = FALSE}
# remove irrelevant strings
data <- all_data[expMode == "experiment"]
data$MidFile <- gsub("./mid/", "", data$midFile)
data$MidFile <- gsub(".mid", "", data$MidFile)

# name conditions
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", midFile)]$Condition <- "teaching"
data[grepl("_p_", midFile)]$Condition <- "performing"
data[grepl("a_", midFile)]$Skill <- "articulation"
data[grepl("d_", midFile)]$Skill <- "dynamics"

# check correct or incorrect
data$Correct <- 0
data[Condition == "teaching" & rating == "Yes"]$Correct <- 1
data[Condition == "performing" & rating == "No"]$Correct <- 1

all <- data[expMode == "experiment", .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill)]
all[order(subjectNumber, Condition, Skill)]
```

## 1. Summary
### All
```{r acc, echo = FALSE}
# individual
ind_data <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

@@

### Articulation vs. Dynamics

```{r plot2, echo = FALSE}
ind_data_ad <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Skill)]
ind_data_ad <- ind_data_ad[order(subjectNumber, Skill)]

subset_ad <- ind_data_ad[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/48)*100, SD_Percent = (sd(Sum)/48)*100), by = .(Skill)]
subset_ad

data_box <- data.table(subjectNumber = ind_data_ad$subjectNumber, articulation = ind_data_ad[Skill == "articulation"]$Percent, dynamics = ind_data_ad[Skill == "dynamics"]$Percent)
box <- ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box
```

## One-sample *t*-test
### Articulation
```{r t-test-a, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "articulation"]$Percent)

t.test(ind_data_ad[Skill == "articulation"]$Percent, mu = 50, alternative = "two.sided")
```
- Not significantly different from chance (p > .2).

### Dynamics
```{r t-test-d, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent)

t.test(ind_data_ad[Skill == "dynamics"]$Percent, mu = 50, alternative = "two.sided")
```

## Paired *t*-test between articulation and dynamics
```{r p-t-test, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent-ind_data_ad[Skill == "articulation"]$Percent)

t.test(ind_data_ad[Skill == "articulation"]$Percent, ind_data_ad[Skill == "dynamics"]$Percent, paired = TRUE)
```

## 2. Teaching stimuli
### All
```{r item1, echo = FALSE}
# individual
ind_data2 <- data[subjectNumber != 9 & Condition == "teaching", .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill, MidFile)]
ind_data2 <- ind_data3[order(subjectNumber, Condition, Skill, MidFile)]

# all
subset2 <- ind_data2[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(Condition, Skill, MidFile)]
subset2[order(-Mean_Percent)]
```

### articulation
```{r item1-a, echo = FALSE}
subset2_a <- subset2[Skill == "articulation"]
subset2_a$MidFile <- gsub("a_t_", "", subset2_a$MidFile)
subset2_a[order(-Mean_Percent)]

bar_a <- ggplot(subset2_a, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar_a
```

### dynamics
```{r item1-d, echo = FALSE}
subset2_d <- subset2[Skill == "dynamics"]
subset2_d$MidFile <- gsub("d_t_", "", subset2_d$MidFile)
subset2_d[order(-Mean_Percent)]

bar_d <- ggplot(subset2_d, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar_d
```

## 3. performing (each item)
### all
```{r item2, echo = FALSE}
# individual
ind_data3 <- data[subjectNumber != 9 & Condition == "performing", .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill, MidFile)]
ind_data3 <- ind_data3[order(subjectNumber, Condition, Skill, MidFile)]

# all
subset3 <- ind_data3[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(Condition, Skill, MidFile)]
subset3[order(-Mean_Percent)]
```

### articulation
```{r item2-a, echo = FALSE}
subset3_a <- subset3[Skill == "articulation"]
subset3_a$MidFile <- gsub("a_p_", "", subset3_a$MidFile)
subset3_a[order(-Mean_Percent)]

bar2_a <- ggplot(subset3_a, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar2_a
```

### dynamics
```{r item2-d, echo = FALSE}
subset3_d <- subset3[Skill == "dynamics"]
subset3_d$MidFile <- gsub("d_p_", "", subset3_d$MidFile)
subset3_d[order(-Mean_Percent)]

bar2_d <- ggplot(subset3_d, aes(x = reorder(MidFile, Mean_Percent), y = Mean_Percent)) + geom_bar(stat = "identity") + labs(x = "Stimuli", y = "Correct %")
bar2_d
```

```{r export, include = FALSE}
order_data <- rbind(subset2, subset3)
